---
title: "readme"
#author: "Paul Q. Sims"
#date: "7/13/2020"
output: 
  github_document
---

- Last update: `r Sys.Date()`

## Description

Data and code associated with the study “An investigation of population variation in maze exploration and its predictors in wild Trinidadian guppies (*Poecilia reticulata*)”. *In prep*. Please see the publication for further details on data collection, the experimental apparatus, procedure, and statistical analyses.

This repository will be permanently archived at Zenodo. 

A report summary containing all the R script output, including plots, can be found at [analysis-summary.md](https://github.com/paulqsims/inno_pop/blob/master/reports/analysis-summary.md) in `reports/`

R Markdown files can be rendered by sourcing [render-Rmd.R](https://github.com/paulqsims/inno_pop/blob/master/R/render-Rmd.R) in `R/`

## Authors

Paul Q. Sims (<paul.q.sims@gmail.com>)

Simon M. Reader (<simon.reader@mcgill.ca>)

## Directory structure

* `data/` Data used in analyses
* `R/` Self-contained R markdown (*.Rmd) scripts (can be run on their own) and functions used for the analyses and in generating plots
* `figs/` High quality figures generated for publication
* `reports/` Markdown and html summaries of code output

## Prerequisites: R version and R packages

`r R.Version()$version.string`

```{r package names, echo=FALSE,warning=FALSE,eval=TRUE,message=FALSE}
# Package dataframe
dat_packages <- setNames(data.frame(matrix(ncol = 3, nrow = 12)),
                        c("Package_Name", "Version", "Function"))
# Package name
dat_packages$Package_Name <- 
  c("tidyverse", "broom", "broom.mixed", "nlme", "MuMIn", "ggplot2",
  "ggsignif", "scales", "effects", "rprojroot", "rmarkdown", "knitr")
# Package function
dat_packages$Function <- 
  c(rep("Data clean up", times = 3), 
    rep("Analyses", time = 2),
    rep("Plots", times = 4),
    rep("Miscellaneous", times = 3))
# Package version
for (i in 1:nrow(dat_packages)) {
   dat_packages$Version[i] <- as.character(packageVersion(dat_packages$Package_Name[i]))
}
knitr::kable(dat_packages)
```

## Metadata

Reference file: [data_Sims-Reader_2020.csv](https://github.com/paulqsims/inno_pop/blob/master/data/data_Sims-Reader_2020.csv)

```{r, echo = FALSE, results = TRUE, message = FALSE}

library(tidyverse)

data_analysis <- 
  read_csv("data/data_Sims-Reader_2020.csv")

# Write out descriptions associated with each column header
var_names <- names(data_analysis)
var_descrip <- c(
  "Population: Lower or upper Aripo",
  "Site sampled",
  "Unique site identifier",
  "Trial: 1 or 2",
  "Group id",
  "Latency to reach the goal zone, zone 11, (sec)",
  "Total zones entered",
  "Body length (mm)",
  "Improvement ratio learning measure = goal zone latency on trial 2/goal zone latency on trial 1")

var_metadata <- tibble(var_names,var_descrip) %>%
                  rename(Column = "var_names", Description = "var_descrip")

knitr::kable(var_metadata)
```

- Missing values are input as `NA` text

## Licenses

Software and code is licensed under the MIT License, data is licensed under CC0, and media files are under CC-BY-4.0 - see the [license.md](https://github.com/paulqsims/inno_pop/blob/master/license.md) file for details.

## Funding information

This work was supported by the Natural Sciences and Engineering Research Council of Canada (NSERC; Discovery grants #418342-2012 and 429385-2012) and the Canada Foundation for Innovation (grant #29433).

