---
title: "data_readme"
output:
  html_document: default
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description

Author: Paul Q. Sims (<paul.q.sims@gmail.com>)

Data associated with the study “An investigation of population variation in maze exploration and its predictors in wild Trinidadian guppies (*Poecilia reticulata*)”. *In prep*. Please see the publication for further details on data collection, the experimental apparatus, procedure, and statistical analyses.

# Metadata

Reference file: [data_Sims-Reader_2020.csv](https://github.com/paulqsims/inno_pop/blob/master/data/data_Sims-Reader_2020.csv)

```{r, echo = FALSE, results = TRUE, message = FALSE}
library(tidyverse)

data_analysis <- 
  read_csv("data/data_Sims-Reader_2020.csv")

# Write out descriptions associated with each column header
var_names <- names(data_analysis)
var_descrip <- c(
  "Population: Lower or upper Aripo",
  "Site sampled",
  "Unique site identifier",
  "Trial: 1 or 2",
  "Group id",
  "Latency to reach the goal zone, zone 11, (sec)",
  "Total zones entered",
  "Body length (mm)")

var_metadata <- tibble(var_names,var_descrip) %>%
                  rename(Column = "var_names", Description = "var_descrip")

knitr::kable(var_metadata)
```

- Missing values are input as `NA` text
