# Analysis: Predictors of innovation 
- Author: Paul Q. Sims
- Contact: paul.q.sims@gmail.com
- Date: 2020
- Purpose: Predictors of innovation analyses for Sims and Reader 2020

```{r inno-predict-knitr-opts}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file(),
                     eval = TRUE, echo = TRUE, message = FALSE,
                     warning = FALSE)
knitr::opts_chunk$set(root.dir = rprojroot::find_rstudio_root_file(),
                     eval = TRUE, echo = TRUE, message = FALSE,
                     warning = FALSE)
```

## Setup

```{r inno-predict-setup}
# Load libraries
library(tidyverse)  # for cleaning and modifying data
library(nlme)  # for mixed models and generalized least squares

# Load personalized functions
source("R/custom-functions.R")

# Read in data
data_analysis <-
  read_mod_data("data_Sims-Reader_2020")

# Create dataset without NAs - lme won't remove them
data_analysis_NA_inno <-
  data_analysis %>%
  select(goal_z_lat_LN, tot_z_sc, pop, group,
         site_uni, body_length_sc, trial) %>%
  drop_na() 
```

## Random-effect testing

Check if there are significant among-group differences

```{r inno-predict-ranef-test}
# Model formula for fixed-effects
inno_full_pred <- formula(goal_z_lat_LN ~ tot_z_sc * pop +  # fixed effect formula
                            body_length_sc * pop +
                            trial * pop)

m_inno_predict_full <-  # Fit model with group
  lme(inno_full_pred,
      weights = varIdent(form = ~ 1|site_uni * pop),
      random = ~ 1 | group,
      data = data_analysis_NA_inno,
      method = "REML")

m1_inno_predict_red_rand <-  # Fit model without group
  gls(inno_full_pred,
      weights = varIdent(form = ~ 1|site_uni * pop),
      data = data_analysis_NA_inno,
      method = "REML")

anova(m_inno_predict_full, m1_inno_predict_red_rand)  # Likelihood ratio test
```

## Fixed-effect selection

Test significance of interactions

```{r inno-predict-fixef-select}
m1_inno_predict_red_fix <-  # Fit full model with ML 
  update(m_inno_predict_full, method = "ML")

# Find largest non-significant p-value for interaction
model_sel_temp1 <- drop1(m1_inno_predict_red_fix, test = "Chi")
model_sel_temp1

# Remove largest non-significant p-value for interaction and update model and continue process
m1_temp <- update(m1_inno_predict_red_fix, ~ . -pop:body_length_sc)  # Remove most non-sig interactions
model_sel_temp2 <- drop1(m1_temp, test = "Chi")  # Update model and check remaining sig interactions
model_sel_temp2

m2_temp <- update(m1_temp, ~ . -pop:trial)  # same as above
model_sel_temp3 <- drop1(m2_temp, test = "Chi")  # same as above
model_sel_temp3

# Final model of innovation predictors
inno_pred_reduc <- update(inno_full_pred, ~ .  # Update model formula
                            -pop:body_length_sc -pop:trial)
```

## Final model of predictors of innovation

Fit final model

```{r inno-predict-fit-final-mod}
# Relevel population for testing total zones slope in other population
# data_analysis_NA_inno$pop <- relevel(data_analysis_NA_inno$pop, "Upper Aripo")  # Upper Aripo baseline
# data_analysis_NA_inno$pop <- relevel(data_analysis_NA_inno$pop, "Lower Aripo")  # Lower Aripo baseline, original 

# Change trial contrasts in order to get marginal effects for average trial 
contrasts(data_analysis_NA_inno$trial) <- c(-1,1)
#contrasts(data_analysis_NA_inno$trial) # check

# For changing trial back to dummy coding
# contrasts(data_analysis_NA_inno$trial) <- c(0,1)
# contrasts(data_analysis_NA_inno$trial) # check

# Fit final reduced model
m_inno_predict_reduc <- 
  lme(inno_pred_reduc,
      weights = varIdent(form = ~ 1|site_uni * pop),
      random = ~ 1 | group,
      data = data_analysis_NA_inno)

# Model summary
summary(m_inno_predict_reduc)

# R2
MuMIn::r.squaredGLMM(m_inno_predict_reduc)
```

Tidy final model summary

```{r inno-predict-finalmod-tidy}
m_inno_predict_final_tidy <-
  m_inno_predict_reduc %>%
  broom.mixed::tidy() %>%
  filter(effect == "fixed") %>%
  mutate(across(.cols = c(estimate:statistic), ~round_est(.x)),
         p_value = round_pval(p.value)) %>%
  select("term", "estimate","std.error", "statistic", "p_value")

m_inno_predict_final_tidy
```

