# Analysis: Population comparisons
- Author: Paul Q. Sims
- Contact: paul.q.sims@gmail.com
- Date: 2020
- Purpose: Population comparison analyses for Sims and Reader 2020

```{r pop-comp-knitr-opts}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file(),
                     eval = TRUE, echo = TRUE, message = FALSE,
                     warning = FALSE)
knitr::opts_chunk$set(root.dir = rprojroot::find_rstudio_root_file(),
                     eval = TRUE, echo = TRUE, message = FALSE,
                     warning = FALSE)
```

## Setup

```{r pop-comp-setup}
# Load libraries
library(tidyverse)  # for cleaning and modifying data
library(nlme)  # for mixed models and generalized least squares
library(broom)  # for tidying model output

# Load personalized functions
source("R/custom-functions.R")

# Read in data
data_analysis <-
  read_mod_data("data_Sims-Reader_2020")
```


## Population comparisons

### Body length, total zones entered

Create list of models for both variables

```{r pop-comp_length-zones}
m_length_totz_popcomp_tidy <- 
  data_analysis %>%
  select(group, pop, trial, body_length_LN, tot_z_LN) %>%
  filter(trial == 1) %>% 
  pivot_longer(cols = c("body_length_LN", "tot_z_LN"),  # reshape data
               names_to = "response",
               values_to = "value") %>%
  group_by(response) %>%
  nest() %>%
  mutate(model = map(data, ~lm(value ~ pop, data = .)),  # run models for each
         tidy_model = map(model, tidy)) %>%  # tidy output
  unnest(tidy_model) %>%
  mutate(across(.cols = c(estimate:statistic), ~ round_est(.x)),  # round values
         p_value = round_pval(p.value)) %>%  
  select(-c(data, model, p.value))  # extract relevant columns
```

Body length population comparison model summary

```{r pop-comp_length-sum}
data_analysis %>%
filter(trial == 1) %>%
lm(body_length_LN ~ pop, data = .) %>%
summary(.)
```

Total zones entered population comparison model summary

```{r pop-comp_totz-sum}
data_analysis %>%
filter(trial == 1) %>%
lm(tot_z_LN ~ pop, data = .) %>%
summary(.)
```

### Innovation: Goal zone latency

Create a data set without NAs - `lme()` won't remove them

```{r}
data_analysis_NA_inno <-
  data_analysis %>%
  select(goal_z_lat_LN, tot_z_sc, pop, group,
         site_uni, body_length_sc, trial) %>%
  drop_na() 
```

Reduced innovation predictor model (see analysis_inno-predict.Rmd for how it was obtained)

```{r}
# Relevel population for testing the slope of total zones entered in other population
# data_analysis_NA$pop <- relevel(data_analysis_NA$pop, "Upper Aripo")  # Upper Aripo baseline
# data_analysis_NA$pop <- relevel(data_analysis_NA$pop, "Lower Aripo")  # Lower Aripo baseline, original baseline

# Change trial contrasts in order to get marginal effects for average trial 
contrasts(data_analysis_NA_inno$trial) <- c(-1,1)
# contrasts(data_analysis_NA_inno$trial) # check

# For changing trial back to dummy coding
# contrasts(mydata$trial.F) <- c(0,1)
# contrasts(mydata$trial.F) # check

# Fit reduced model
m_inno_predict_final <- 
  lme(goal_z_lat_LN ~ tot_z_sc * pop + 
        body_length_sc + trial,
      weights = varIdent(form = ~ 1|site_uni * pop),
      random = ~ 1 | group,
      data = data_analysis_NA_inno,
      method = "REML")

# Tidy innovation predictor model
m_inno_predict_final_tidy <-
  m_inno_predict_final %>%
  broom.mixed::tidy() %>%
  filter(effect == "fixed",
         term %in% c("(Intercept)", "popUpper Aripo")) %>%
  mutate(response = "goal_z_lat_LN",
         across(.cols = c(estimate:statistic), ~round_est(.x)),
         p_value = round_pval(p.value)) %>%
  select("response", "term", "estimate","std.error", "statistic", "p_value") 
```

### Learning: Improvement ratio

Create dataset without NAs - `lme()` won't remove them

```{r}
data_analysis_NA_learn <-
  data_analysis %>%
  filter(trial == 1) %>%
  select(learn_prop_LN, tot_z_sc, pop, group,
         site_uni, body_length_sc) %>%
  drop_na() 
```

Build full model

```{r}
m_learn_pop_comp <- 
  gls(learn_prop_LN ~ pop,
      weights = varIdent(form = ~ 1|site_uni * pop),
      data = data_analysis_NA_learn,
      method = "REML")
```

Population differences in learning model summary

```{r}
summary(m_learn_pop_comp)
```

Reduced learning model (no population differences)

```{r}
m_learn <- update(m_learn_pop_comp, ~ 1)  # update model to intercept only

# Model summary: learning overall
summary(m_learn)

# Tidy learning population comparison model
m_learn_pop_comp_tidy <-
  m_learn_pop_comp %>%
  broom.mixed::tidy() %>%
  mutate(response = "improvement_ratio",
         across(.cols = c(estimate:statistic), ~round_est(.x)),
         p_value = round_pval(p.value)) %>%
  select("response", "term", "estimate","std.error", "statistic", "p_value") 
```

## Cleanup and final results

```{r}
# Bind all population comparisons together
m_pop_comp_final_tidy <-
  bind_rows(m_learn_pop_comp_tidy, m_inno_predict_final_tidy,
            m_length_totz_popcomp_tidy) %>%
  rename(predictor = "term")

print(m_pop_comp_final_tidy)  # See reduced inno predictor model for trial measure of learning
```

