# Analysis: Population comparisons
- Author: Paul Q. Sims
- Contact: paul.q.sims@gmail.com
- Date: 2020
- Purpose: Population comparison analyses for Sims and Reader 2020

```{r pop-comp-knitr-opts}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file(),
                     eval = TRUE, echo = TRUE, message = FALSE,
                     warning = FALSE)
knitr::opts_chunk$set(root.dir = rprojroot::find_rstudio_root_file(),
                     eval = TRUE, echo = TRUE, message = FALSE,
                     warning = FALSE)
```

## Setup

```{r pop-comp-setup}
# Load libraries
library(tidyverse)  # for cleaning and modifying data
library(nlme)  # for mixed models and generalized least squares
library(broom)  # for tidying model output

# Load personalized functions
source("R/custom-functions.R")

# Read in data
data_analysis <-
  read_mod_data("data_Sims-Reader_2020")
```

## Population comparisons

### Body length, total zones entered

Create list of models for both variables

```{r pop-comp_length-zones}
m_length_totz_popcomp_tidy <- 
  data_analysis %>%
  select(group, pop, trial, body_length_LN, tot_z_LN) %>%
  filter(trial == 1) %>% 
  pivot_longer(cols = c("body_length_LN", "tot_z_LN"),  # reshape data
               names_to = "response",
               values_to = "value") %>%
  group_by(response) %>%
  nest() %>%
  mutate(model = map(data, ~lm(value ~ pop, data = .)),  # run models for each
         tidy_model = map(model, tidy)) %>%  # tidy output
  unnest(tidy_model) %>%
  mutate(across(.cols = c(estimate:statistic), ~ round_est(.x)),  # round values
         p.value = round_pval(p.value)) %>%  
  select(-c(data, model))  # extract relevant columns 
```

Body length population comparison model summary

```{r pop-comp_length-sum}
data_analysis %>%
filter(trial == 1) %>%
lm(body_length_LN ~ pop, data = .) %>%
pretty_PredictTab(.)
```

Total zones entered population comparison model summary

```{r pop-comp_totz-sum}
data_analysis %>%
filter(trial == 1) %>%
lm(tot_z_LN ~ pop, data = .) %>%
pretty_PredictTab(.)
```

### Innovation: Goal zone latency

Create a data set without NAs - `lme()` won't remove them

```{r}
data_analysis_NA_inno <-
  data_analysis %>%
  select(goal_z_lat_LN, pop, group,
         site_uni, trial) %>%
  drop_na() 
```

Fit innovation model

```{r}
# Change trial contrasts in order to get marginal effects for average trial 
contrasts(data_analysis_NA_inno$trial) <- c(-1,1)
# contrasts(data_analysis_NA_inno$trial) # check

# Fit reduced model
m_inno_pop_comp <- 
  lme(goal_z_lat_LN ~ pop + trial,
      weights = varIdent(form = ~ 1|site_uni * pop),
      random = ~ 1 | group,
      data = data_analysis_NA_inno,
      method = "REML")

# Tidy innovation predictor model
m_inno_pop_comp_tidy <-
  m_inno_pop_comp %>%
  pretty_PredictTab(., mixedModel = TRUE, kable = FALSE) %>%
  mutate(response = "goal_zone_lat_LN") %>%
  relocate(response)
```

### Learning: Improvement ratio

Create dataset without NAs - `lme()` won't remove them

```{r}
data_analysis_NA_learn <-
  data_analysis %>%
  filter(trial == 1) %>%
  select(learn_prop_LN, pop, group, site_uni) %>%
  drop_na() 
```

Build full model

```{r}
m_learn_pop_comp <- 
  gls(learn_prop_LN ~ pop,
      weights = varIdent(form = ~ 1|site_uni * pop),
      data = data_analysis_NA_learn,
      method = "REML")
```

Population differences in learning model summary

```{r}
summary(m_learn_pop_comp)

# Tidy learning population comparison model
m_learn_pop_comp_tidy <-
  m_learn_pop_comp %>%
  pretty_PredictTab(., mixedModel = TRUE, kable = FALSE) %>%
  mutate(response = "improvement_ratio") %>%
  relocate(response)
```

Reduced learning model (no population differences)

```{r}
m_learn <- update(m_learn_pop_comp, ~ 1)  # update model to intercept only

# Model summary: learning overall
pretty_PredictTab(m_learn)
```

## Cleanup and final results

```{r}
# Bind all population comparisons together
m_pop_comp_final_tidy <-
  bind_rows(m_learn_pop_comp_tidy, m_inno_pop_comp_tidy,
            m_length_totz_popcomp_tidy) %>%
  rename(predictor = "term")

m_pop_comp_final_tidy  # See reduced inno predictor model for trial measure of learning
```

