# Analysis: Predictors of innovation 
- Author: Paul Q. Sims
- Contact: paul.q.sims@gmail.com
- Date: 2020
- Purpose: Predictors of innovation analyses for Sims and Reader 2020

```{r inno-predict-knitr-opts}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file(),
                     eval = TRUE, echo = TRUE, message = FALSE,
                     warning = FALSE)
knitr::opts_chunk$set(root.dir = rprojroot::find_rstudio_root_file(),
                     eval = TRUE, echo = TRUE, message = FALSE,
                     warning = FALSE)
```

## Setup

```{r inno-predict-setup}
# Load libraries
library(tidyverse)  # for cleaning and modifying data
library(nlme)  # for mixed models and generalized least squares

# Load personalized functions
source("R/custom-functions.R")

# Read in data
data_analysis <-
  read_mod_data("data_Sims-Reader_2020")

# Create dataset without NAs - lme won't remove them
data_analysis_NA_inno <-
  data_analysis %>%
  select(goal_z_lat_LN, tot_z_sc, pop, group,
         site_uni, body_length_sc, trial) %>%
  drop_na() 
```

## Random-effect testing

Check if there are significant among-group differences

```{r inno-predict-ranef-test}
# Model formula for fixed-effects
inno_full_pred <- formula(goal_z_lat_LN ~ tot_z_sc * pop + 
                            body_length_sc * pop +
                            trial * pop)

# Fit model with group
m_inno_predict_full <-  
  lme(inno_full_pred,
      weights = varIdent(form = ~ 1|site_uni * pop),
      random = ~ 1 | group,
      data = data_analysis_NA_inno,
      method = "REML")

# Fit model without group
m1_inno_predict_red_rand <-  
  gls(inno_full_pred,
      weights = varIdent(form = ~ 1|site_uni * pop),
      data = data_analysis_NA_inno,
      method = "REML")

# Likelihood ratio test
anova(m_inno_predict_full, m1_inno_predict_red_rand) %>%
  as_tibble(.) %>%
  mutate(call = str_trunc(as.character(.$call), width = 10)) %>%  # condense call
  knitr::kable(., digits = 2, align = "l",
               caption = "Likielihood ratio test for group random effect") 
```

## Fixed-effect selection

Test significance of interactions

```{r inno-predict-fixef-select}
# Fit full model with ML 
m1_inno_predict_red_fix <-  
  update(m_inno_predict_full, method = "ML")

# Find largest non-significant p-value for interaction
model_sel_temp1 <- drop1(m1_inno_predict_red_fix, test = "Chi")
model_sel_temp1 %>%
  rd_stepwise_out(.) %>%
  knitr::kable(.) 

# Remove largest non-significant p-value for interaction and update model and continue process
m1_temp <- update(m1_inno_predict_red_fix, ~ . -pop:body_length_sc)  # Remove most non-sig interactions
model_sel_temp2 <- drop1(m1_temp, test = "Chi")  # Update model and check remaining sig interactions
model_sel_temp2 %>%
  rd_stepwise_out(.) %>%
  knitr::kable(., caption = "No evidence for learning (i.e. trial) differences between populations (pop:trial)")

m2_temp <- update(m1_temp, ~ . -pop:trial)  # same as above
model_sel_temp3 <- drop1(m2_temp, test = "Chi")  # same as above
model_sel_temp3 %>%
  rd_stepwise_out(.) %>%
  knitr::kable(.)

# Final model of innovation predictors
inno_pred_reduc <- update(inno_full_pred, ~ .  # Update fixef model formula
                            -pop:body_length_sc -pop:trial)
```

## Final model of predictors of innovation

Fit final model

```{r inno-predict-fit-final-mod}
# Relevel population for testing total zones slope in other population
# data_analysis_NA_inno$pop <- relevel(data_analysis_NA_inno$pop, "Upper Aripo")  # Upper Aripo baseline
# data_analysis_NA_inno$pop <- relevel(data_analysis_NA_inno$pop, "Lower Aripo")  # Lower Aripo baseline, original 

# Fit final reduced model
m_inno_predict_reduc <- 
  lme(inno_pred_reduc,
      weights = varIdent(form = ~ 1|site_uni * pop),
      random = ~ 1 | group,
      contrasts = list(trial = c(-1,1)),  # Change trial contrasts in order to get marginal effects for average trial 
      data = data_analysis_NA_inno,
      method = "REML")
```

Final model summary

```{r inno-predict-finalmod-tidy}
# Tidy model output
pretty_PredictTab(m_inno_predict_reduc,
                  title = "Predictors of goal zone latency",
                  mixedModel = TRUE) 

# R squared (delta)
knitr::kable(MuMIn::r.squaredGLMM(m_inno_predict_reduc),
      digits = 2, align = "l",
      caption = "Marginal and conditional R2")
```

